R for Clinical Study Reports and Submission

### 2021-11-09 project-folder.Rmd

We illustrate the R package project folder for the A&R project using a demo project called
[`esubdemo`](https://github.com/elong0527/esubdemo.git).
You can
[clone](https://happygitwithr.com/rstudio-git-github.html#clone-the-new-github-repository-to-your-computer-via-rstudio)
the project using RStudio.

@@

> Under the `vignettes` folder, there are two folders: `adam` folder and `tlf` folder.
> The `adam` folder contains ADaM datasets.
> The `tlf` folder contains output TLFs in RTF format.
> We put `adam` and `tlf` folders within the `vignettes` folder only for illustration purposes.
> In an actual A&R report, you may have a different location to save your input and output.

```
vignettes
├── adam_data
├── tlf
├── tlf-01-disposition.Rmd
├── tlf-02-population.Rmd
├── tlf-03-baseline.Rmd
├── tlf-04-efficacy.Rmd
├── tlf-05-ae-summary.Rmd
└── tlf-06-ae-spec.Rmd
```

@@

> This is the same level of reproducibility in most SAS environments.
> https://support.sas.com/techsup/pcn/altopsys.html

@@

If there is a R version mismatch, an error message is displayed as below.

@@

> `.Rprofile` is only for project-specific R packages.
> A standard R package should not use `.Rprofile`.

@@

> RStudio Package Manager (RSPM) provided a solution to host both
> publicly available and internally developed R packages.
> However, the public RSPM server (https://packagemanager.rstudio.com/client/#/repos/2/overview)
> did not provide a daily snapshot as MRAN had.

@@

- create a `renv.lock` file to save the state of package versions.
- create a `renv` folder to manage R packages for a project.

> The `renv.lock` file and `renv/` folder are only for project-specific R package.
> A standard R package should not use `renv`.

@@

If the project is initiated properly,
you shall be able to see similar messages to inform how we control R package versions.

@@

> A cloud-based R environment (e.g., RStudio Workbench)
> can enhance the reproducibility within an organization
> by using the same operating system, R version, and R package versions for an A&R project.
> More details can be found at <https://environments.rstudio.com/>.

> A container solution like [Docker](https://github.com/rstudio/r-docker) could further enhance
> the reproducibility across an organization at the operating system level but beyond the scope of this book.

@@

- Create project template using [RStudio Project Templates](https://rstudio.github.io/rstudio-extensions/rstudio_project_templates.html)
- Add additional compliance checks for analysis and reporting.
- Save log files for running in batch.

>>>>

We will use the [`esubdemo`](https://github.com/elong0527/esubdemo) R package to
illustrate the project folder structure for the A&R project. You can
[clone](https://happygitwithr.com/rstudio-git-github.html#clone-the-new-github-repository-to-your-computer-via-rstudio)
the project using RStudio with

```bash
git clone https://github.com/elong0527/esubdemo.git
```

@@

::: {.rmdnote}
Under the `vignettes/` folder, there are two folders: `adam/` and `tlf/`.
The `adam/` folder contains ADaM datasets.
The `tlf/` folder contains output TLFs in RTF format.
We put `adam/` and `tlf/` folders within the `vignettes/` folder only for illustration purposes.
In an actual A&R report, you may have a different location to save your input and output.
:::

```
vignettes
├── data-adam
├── tlf
├── tlf-01-disposition.Rmd
├── tlf-02-population.Rmd
├── tlf-03-baseline.Rmd
├── tlf-04-efficacy.Rmd
├── tlf-05-ae-summary.Rmd
└── tlf-06-ae-spec.Rmd
```

@@

::: {.rmdnote}
This is the same level of reproducibility in most SAS environments:
<https://support.sas.com/techsup/pcn/altopsys.html>
:::

### R version

@@

If there is an R version mismatch, an error message is displayed as below.

@@

::: {.rmdnote}
`.Rprofile` is only for project-specific R packages.
A standard R package should not use `.Rprofile`.
:::

### R package version

@@

::: {.rmdnote}
RStudio Package Manager (RSPM) provides a solution to host both
publicly available and internally developed R packages.
However, the [public RSPM server](https://packagemanager.rstudio.com/client/#/repos/2/overview)
does not provide a daily snapshot as MRAN does.
:::

@@

- create a `renv.lock` file to save the state of package versions.
- create a `renv/` folder to manage R packages for a project.

::: {.rmdnote}
The `renv.lock` file and `renv/` folder are only for project-specific R package.
A standard R package should not use `renv`.
:::

@@

If the project is initiated properly,
you should be able to see similar messages to inform how we control R package versions.

@@

::: {.rmdnote}
A cloud-based R environment (e.g., RStudio Workbench)
can enhance the reproducibility within an organization
by using the same operating system, R version, and R package versions for an A&R project.
More details can be found at <https://environments.rstudio.com/>.
:::

::: {.rmdnote}
A container solution like Docker [@RJ-2020-007] could further enhance
the reproducibility across an organization at the operating system level
but beyond the scope of this book.
:::

@@

- Create project template using
  [RStudio project templates](https://rstudio.github.io/rstudio-extensions/rstudio_project_templates.html);
- Add additional compliance checks for analysis and reporting;
- Save log files for running in batch.
